---
title: "Mitochondial mutational spectrum of cancer as a function of position (TBSS), tissue turn overrate  and VAF"
author: "Konstantin Popadin"
date: "11/05/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls(all=TRUE))
library(knitr)
library(tidyverse)
library(kableExtra)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lsa) # install.packages("lsa"); install.packages("SnowballC")
```
## Background
<br>
Cancer paper:
Dataset is 7611 somatic mtDNA mutations
<br>
Hypothesis: A>G is a function of TBSS, TissueTurnOver and VAF
<br><br>

#### 1. Read a cancer dataset and describe key variables
***
```{r}
Mut = read.table("../data/1raw/mtDNA_snv_Oct2016.txt", head = TRUE, sep = '\t')  
names(Mut)
Mut$Subst = paste(Mut$ref,Mut$var,sep = '>')
table(Mut$Subst) # light chain notation is be default
table(Mut$Tier2) # cancer tissue
Mut$tumor_var_freq = as.numeric(gsub('%','',Mut$tumor_var_freq))
summary(Mut$tumor_var_freq) # 0.01    1.68    4.37   19.96   25.39   99.86
hist(Mut$position, breaks = 200)
Mut$DummyHighVaf = 0
for (i in 1:nrow(Mut))
{
  if (Mut$tumor_var_freq[i] >= 25.39) # 4.37
  {Mut$DummyHighVaf[i] = 1}  
}
table(Mut$DummyHighVaf)
```
<br>
There are many mutations from D loop [16024 - 576 according to MitoWheel]. 
Since they mutate according to different rules, it is better to delete all of them in the analyses.
In this case for all other mutations we can derive time being single stranded.
<br><br>

#### 2. Create a dataset with tissue-specific turnover rates and merge it to the Mut
***
```{r}
CancerTissue = c('Bladder','Bone/SoftTissue','Breast','Biliary','Cervix','Lymphoid','Myeloid','Colon/Rectum','Prostate','Esophagus','Stomach','CNS','Head/Neck','Kidney','Liver','Lung','Ovary','Pancreas','Skin','Thyroid','Uterus')  
TurnOverDays = c(200,5373,84.5,200,6,30,30,5,120,11,5.5,10000,16,1000,400,5143,11000,360,147,4138,4); length(TurnOverDays)
Turn = data.frame(CancerTissue,TurnOverDays)
Turn = Turn[order(Turn$TurnOverDays),]
summary(Turn$TurnOverDays) #  4      16     147    1823    1000   11000
Turn$DummyNotFastTurnOver = 0; # !!!! 
Turn$DummySlowTurnOver = 0;
Turn$DummyVerySlowTurnOver = 0;
for (i in 1:nrow(Turn)) 
{ 
if (Turn$TurnOverDays[i] > 1000) {Turn$DummyVerySlowTurnOver[i] = 1} # if  slower than the third quartile (1000 days)
if (Turn$TurnOverDays[i] > 147)  {Turn$DummySlowTurnOver[i] = 1}  # if  slower than the median (147 days) median in Mut == 360
if (Turn$TurnOverDays[i] > 16)  {Turn$DummyNotFastTurnOver[i] = 1}  # if  slower than the median (147 days)
}

Mut = merge(Mut,Turn, by.x = 'Tier2', by.y = 'CancerTissue')
summary(Mut$TurnOverDays) # there are no NA - good
```
<br><br>

#### 3. Derive TBSS and DummyHighTbss for each position, positions within D-loop and Ol will have NA
***
```{r}
Mut$TBSS = NA
for (i in 1:nrow(Mut))
{ # i = 1
  if (Mut$position[i] > 576 & Mut$position[i] < 5721) # minor arc
    {Mut$TBSS[i] = 16569-(5721 - Mut$position[i])*2} # L-strand replication origin (locus MT-OLR, 5721 - 5798)
  if (Mut$position[i] > 5798 & Mut$position[i] < 16024) # major arc
    {Mut$TBSS[i] = (Mut$position[i] - 5798)*2}  
}
summary(Mut$TBSS)

Mut$DummyHighTbss = NA
for (i in 1:nrow(Mut))
{ # i = 1
  if (!is.na(Mut$TBSS[i]))
  {
  if (Mut$TBSS[i] >= median(Mut[!is.na(Mut$TBSS),]$TBSS)) # 11189
  {Mut$DummyHighTbss[i] = 1}
  if (Mut$TBSS[i] < median(Mut[!is.na(Mut$TBSS),]$TBSS)) # 11189
  {Mut$DummyHighTbss[i] = 0}
  }
}
summary(Mut$DummyHighTbss)
table(Mut$DummyHighTbss)


plot(Mut$position,Mut$TBSS) 
```
<br><br>

### 4. add several dummy variables for logistic regression: Ah>Gh, Ch>Th, Ts, Tv
***
```{r}
Mut$DummyAhGh = 0; Mut$DummyChTh = 0; Mut$DummyTs = 0; Mut$DummyTv = 0;
for (i in 1:nrow(Mut))
{ # i=1
  if (Mut$Subst[i] == 'T>C') {Mut$DummyAhGh[i] = 1}
  if (Mut$Subst[i] == 'G>A') {Mut$DummyChTh[i] = 1}
  if (Mut$Subst[i] %in% c('G>A','T>C','C>T','A>G')) {Mut$DummyTs[i] = 1}
  if (!Mut$Subst[i] %in% c('G>A','T>C','C>T','A>G')) {Mut$DummyTv[i] = 1}
}
table(Mut$DummyAhGh)
table(Mut$DummyChTh)
table(Mut$DummyTs)
table(Mut$DummyTv)
```
<br><br>

### 5. univariate analyses: DummyAhGh ~ TBSS, DummyAhGh ~ + TurnOverDays(*), DummyAhGh ~ + tumor_var_freq(***)
***
```{r}

### AhGh ~ TBSS (logistic regression, Mann-Whitney U test);
### AhGh ~ DummyHighTbss (fisher test)

summary(glm(Mut$DummyAhGh ~ Mut$TBSS, family = binomial)) # 0.0158 and positive
summary(glm(Mut[!is.na(Mut$TBSS),]$DummyAhGh ~ Mut[!is.na(Mut$TBSS),]$TBSS, family = binomial)) # 0.0158 and positive

wilcox.test(Mut[Mut$DummyAhGh == 0,]$TBSS,Mut[Mut$DummyAhGh == 1,]$TBSS) # significant
boxplot(Mut[Mut$DummyAhGh == 0,]$TBSS,Mut[Mut$DummyAhGh == 1,]$TBSS, notch = TRUE) # significant

AgInHighTbss = nrow(Mut[Mut$DummyAhGh == 1 & !is.na(Mut$DummyHighTbss) & Mut$DummyHighTbss == 1,])
AgInLowTbss = nrow(Mut[Mut$DummyAhGh == 1 & !is.na(Mut$DummyHighTbss) & Mut$DummyHighTbss == 0,])
NotAgInHighTbss = nrow(Mut[Mut$DummyAhGh == 0 & !is.na(Mut$DummyHighTbss) & Mut$DummyHighTbss == 1,])
NotAgInLowTbss = nrow(Mut[Mut$DummyAhGh == 0 & !is.na(Mut$DummyHighTbss) & Mut$DummyHighTbss == 0,])

X = cbind(c(AgInHighTbss,AgInLowTbss),c(NotAgInHighTbss,NotAgInLowTbss))
fisher.test(X) # odds = 1.19, p = 0.0014
mosaicplot(X)

summary(glm(Mut$DummyChTh ~ Mut$TBSS, family = binomial)) # 0.06 

### AhGh ~ TurnOver (logistic)
### AhGh ~ DummyNotFastTurnOver (> lower quartile)! We need to explain - why we took this metrique! 

summary(glm(Mut$DummyAhGh ~ Mut$TurnOverDays, family = binomial)) # 0.0377 and positive
summary(glm(Mut[!is.na(Mut$TBSS),]$DummyAhGh ~ Mut[!is.na(Mut$TBSS),]$TurnOverDays, family = binomial))

wilcox.test(Mut[Mut$DummyAhGh == 1,]$TurnOverDays,Mut[Mut$DummyAhGh == 0,]$TurnOverDays) # nothing

Turn$DummySlowTurnOver

AgInSlowTurnOver = nrow(Mut[Mut$DummyAhGh == 1 & Mut$DummyNotFastTurnOver == 1,])
AgInHighTurnOver = nrow(Mut[Mut$DummyAhGh == 1 & Mut$DummyNotFastTurnOver == 0,])
NotAgInSlowTurnOver = nrow(Mut[Mut$DummyAhGh == 0 & Mut$DummyNotFastTurnOver == 1,])
NotAgInHighTurnOver = nrow(Mut[Mut$DummyAhGh == 0 & Mut$DummyNotFastTurnOver == 0,])

X = cbind(c(AgInSlowTurnOver,AgInHighTurnOver),c(NotAgInSlowTurnOver,NotAgInHighTurnOver))
fisher.test(X) # odds = 1.25, p = 0.002
mosaicplot(X)

summary(glm(Mut$DummyChTh ~ Mut$TurnOverDays, family = binomial)) # 0.6
summary(glm(Mut[!is.na(Mut$TBSS),]$DummyChTh ~ Mut[!is.na(Mut$TBSS),]$TurnOverDays, family = binomial))

### AhGh ~ VAF (logistic)
### AhGh ~ VAF Dummy (median and lower quartile were not siginficant, took upper quartile)

summary(glm(Mut$DummyAhGh ~ Mut$tumor_var_freq, family = binomial))
summary(glm(Mut[!is.na(Mut$TBSS),]$DummyAhGh ~ Mut[!is.na(Mut$TBSS),]$tumor_var_freq, family = binomial))

wilcox.test(Mut[Mut$DummyAhGh == 1,]$tumor_var_freq,Mut[Mut$DummyAhGh == 0,]$tumor_var_freq)
boxplot(Mut[Mut$DummyAhGh == 1,]$tumor_var_freq,Mut[Mut$DummyAhGh == 0,]$tumor_var_freq)

AgInHighVaf = nrow(Mut[Mut$DummyAhGh == 1 & Mut$DummyHighVaf == 1,])
AgInLowVaf = nrow(Mut[Mut$DummyAhGh == 1 & Mut$DummyHighVaf == 0,])
NotAgInHighVaf = nrow(Mut[Mut$DummyAhGh == 0 & Mut$DummyHighVaf == 1,])
NotAgInLowVaf = nrow(Mut[Mut$DummyAhGh == 0 & Mut$DummyHighVaf == 0,])

X = cbind(c(AgInHighVaf,AgInLowVaf),c(NotAgInHighVaf,NotAgInLowVaf))
fisher.test(X) # odds = 1.16, p = 0.0102
mosaicplot(X)

summary(glm(Mut$DummyChTh ~ Mut$tumor_var_freq, family = binomial))
summary(glm(Mut[!is.na(Mut$TBSS),]$DummyChTh ~ Mut[!is.na(Mut$TBSS),]$tumor_var_freq, family = binomial))
```
<br><br>

### 6. multivariate backward stepwise logistic regression: DummyAhGh ~ TBSS + TurnOverDays + tumor_var_freq. We need to choose the best equation (several very similar are below). Interestingly, effect size is rather similar (scaled coefficients) and thus we can sum up votings (1 + 1 + 1) to derive the integer hypoxia level in our cosinus exersises.

***
```{r}
summary(glm(Mut$DummyAhGh ~ Mut$TBSS + Mut$TurnOverDays + Mut$tumor_var_freq, family = binomial)) 
summary(glm(Mut$DummyAhGh ~ scale(Mut$TBSS) + scale(Mut$TurnOverDays) + scale(Mut$tumor_var_freq), family = binomial)) 
summary(glm(Mut[!is.na(Mut$TBSS),]$DummyAhGh ~ Mut[!is.na(Mut$TBSS),]$TBSS + Mut[!is.na(Mut$TBSS),]$TurnOverDays + Mut[!is.na(Mut$TBSS),]$tumor_var_freq, family = binomial)) # the same => glm eliminates all rows with at least one NA

summary(glm(Mut$DummyAhGh ~ Mut$TBSS + Mut$DummyNotFastTurnOver + Mut$tumor_var_freq, family = binomial)) # !!!!!
summary(glm(Mut$DummyAhGh ~ scale(Mut$TBSS) + scale(Mut$DummyNotFastTurnOver) + scale(Mut$tumor_var_freq), family = binomial)) # !!!!!

summary(glm(Mut$DummyAhGh ~ Mut$DummyHighTbss + Mut$DummyNotFastTurnOver + Mut$tumor_var_freq, family = binomial)) # !!
summary(glm(Mut$DummyAhGh ~ scale(Mut$DummyHighTbss) + scale(Mut$DummyNotFastTurnOver) + scale(Mut$tumor_var_freq), family = binomial)) # !!

summary(glm(Mut$DummyAhGh ~ Mut$DummyHighTbss + Mut$DummyNotFastTurnOver + Mut$DummyHighVaf, family = binomial)) # !!
summary(glm(Mut$DummyAhGh ~ scale(Mut$DummyHighTbss) + scale(Mut$DummyNotFastTurnOver) + scale(Mut$DummyHighVaf), family = binomial)) # !!

summary(glm(Mut$DummyAhGh ~ Mut$TurnOverDays + Mut$tumor_var_freq, family = binomial)) 
summary(glm(Mut[!is.na(Mut$TBSS),]$DummyAhGh ~ Mut[!is.na(Mut$TBSS),]$TurnOverDays + Mut[!is.na(Mut$TBSS),]$tumor_var_freq, family = binomial))

summary(glm(Mut$DummyAhGh ~ Mut$DummyNotFastTurnOver + Mut$tumor_var_freq, family = binomial)) # !!!!
summary(glm(Mut[!is.na(Mut$TBSS),]$DummyAhGh ~ Mut[!is.na(Mut$TBSS),]$DummyNotFastTurnOver + Mut[!is.na(Mut$TBSS),]$tumor_var_freq, family = binomial)) # still significant but both p-values become higher
```
<br><br>

### 7. cosinus distance between different cancer types, VAF groups and locations
Cosine similarity is lower between biologically different sets of mutations: hypoxic (fast turnover) versus normoxic (low turnover). We need to compare it with random split of mutations! 
Cosine similarity is lower between biologically different sets of mutations: low VAF (recent and hypoxic) versus high VAF (old and normoxic)

***
```{r}
# derive context for each mutation
for (i in 1:nrow(Mut))
{
Mut$SubstWithContext[i] = paste(unlist(strsplit(Mut$X[i],''))[10],Mut$Subst[i],unlist(strsplit(Mut$X.1[i],''))[1],sep = '')  
}

# derive 192 MutSpec for different groups of mutations (by tissue type, VAF, location):

FunCosineSimOfTwoLists <- function(a,b) 
  {
  A = data.frame(table(a)); names(A) = c('subst','AFreq')
  B = data.frame(table(b)); names(B) = c('subst','BFreq')
  AB = merge(A,B)
  cos = cosine(AB$AFreq, AB$BFreq) # 0.98
  # print (cos)
  }

# TurnOverRate
table(Mut$DummySlowTurnOver)
a = Mut[Mut$DummySlowTurnOver == 1,]$SubstWithContext
b = Mut[Mut$DummySlowTurnOver == 0,]$SubstWithContext
FunCosineSimOfTwoLists(a,b) # 0.985

# VAF
summary(Mut$tumor_var_freq) # 4.37
a = Mut[Mut$tumor_var_freq <= 4.37,]$SubstWithContext
b = Mut[Mut$tumor_var_freq > 4.37,]$SubstWithContext
FunCosineSimOfTwoLists(a,b) # 0.986
a = Mut[Mut$tumor_var_freq <= 1.68,]$SubstWithContext
b = Mut[Mut$tumor_var_freq > 25.39,]$SubstWithContext
Reality = as.numeric(FunCosineSimOfTwoLists(a,b)) # 0.967

PermVec = c()
for (permut in 1:1000)
{
Mut$ReSampled.tumor_var_freq = sample(Mut$tumor_var_freq)
a = Mut[Mut$ReSampled.tumor_var_freq <= 1.68,]$SubstWithContext
b = Mut[Mut$ReSampled.tumor_var_freq > 25.39,]$SubstWithContext
PermVec = c(PermVec,as.numeric(FunCosineSimOfTwoLists(a,b)))
}
length(PermVec)
length(PermVec[PermVec < Reality])/length(PermVec) # p = 0.18


# Slow versus High VAF (don't consider intersected mutations which are in both)
a = Mut[Mut$DummySlowTurnOver == 1,]$SubstWithContext # slow () 
b = Mut[Mut$tumor_var_freq > 4.37,]$SubstWithContext # high VAF
FunCosineSimOfTwoLists(a,b) # 0.9949

# MIXED GROUP (Slow and Recent) VS ANOTHER MIXED GROUP (Fast and Old) 
a = Mut[Mut$DummySlowTurnOver == 1 & Mut$tumor_var_freq <= 4.37,]$SubstWithContext
b = Mut[Mut$DummySlowTurnOver == 0 & Mut$tumor_var_freq > 4.37,]$SubstWithContext
FunCosineSimOfTwoLists(a,b) # 0.978 

# VERY MIXED GROUP (Slow and very Recent) VS ANOTHER VERY MIXED GROUP (Fast and very Old) 
a = Mut[Mut$DummySlowTurnOver == 1 & Mut$tumor_var_freq <= 1.68,]$SubstWithContext
b = Mut[Mut$DummySlowTurnOver == 0 & Mut$tumor_var_freq > 25.39,]$SubstWithContext
FunCosineSimOfTwoLists(a,b) # 0.953 

#  NORMOXIC (Slow and Old) VS HYPOXIC (Fast and Recent)
a = Mut[Mut$DummySlowTurnOver == 1 & Mut$tumor_var_freq > 4.37,]$SubstWithContext
b = Mut[Mut$DummySlowTurnOver == 0 & Mut$tumor_var_freq <= 4.37,]$SubstWithContext
FunCosineSimOfTwoLists(a,b) # 0.968 

# VERY NORMOXIC (Slow and Old) VS VERY HYPOXIC (Fast and Recent)
a = Mut[Mut$DummySlowTurnOver == 1 & Mut$tumor_var_freq > 25.39,]$SubstWithContext
b = Mut[Mut$DummySlowTurnOver == 0 & Mut$tumor_var_freq <= 1.68,]$SubstWithContext
Reality = as.numeric(FunCosineSimOfTwoLists(a,b)) # 0.942 

PermVec = c()
for (permut in 1:1000)
{
Mut$ReSampled.tumor_var_freq = sample(Mut$tumor_var_freq)
Mut$ReSampled.DummySlowTurnOver = sample(Mut$DummySlowTurnOver)
a = Mut[Mut$ReSampled.DummySlowTurnOver == 1 & Mut$ReSampled.tumor_var_freq > 25.39,]$SubstWithContext
b = Mut[Mut$ReSampled.DummySlowTurnOver == 0 & Mut$ReSampled.tumor_var_freq <= 1.68,]$SubstWithContext
PermVec = c(PermVec,as.numeric(FunCosineSimOfTwoLists(a,b)))
}
length(PermVec)
length(PermVec[PermVec < Reality])/length(PermVec) # p = 0.017

# may be we have to use only Ts for cosine similarity?



```

### OOOOOLD CODE - DELETE / CHANGE 
*** 

HypMut$AhGhDummy = 0
for (i in 1:nrow(HypMut)) {if(HypMut$Subst[i] == 'T>C') {HypMut$AhGhDummy[i] = 1}}

#################################################
####### ANALYSES:
#################################################

##### 1:  if tissue-specific hypoxia is associated with turnover rate? YES, there is a trend

Agg = aggregate(HypMut$hypoxia_score_buffa, by = list(HypMut$Tier2,HypMut$TurnOverDays), FUN = median)
names(Agg) = c('CancerTissue','TurnOverDays','MedianHypoxiaScoreBuffa')
plot(Agg$TurnOverDays,Agg$MedianHypoxiaScoreBuffa)
cor.test(Agg$TurnOverDays,Agg$MedianHypoxiaScoreBuffa, method = 'spearman', alternative = 'less') # rho = -0.44, p = 0.02704

##### 2:  if slow- middle- and fast- dividing tissues (as in BioRxiv) have different hypoxia (N = 19, N = 828)?


##### 3A:  if there is correlation between VAF and hypoxia?  # YES!!! advanced cancers (with high VAF) are more hypoxic 
cor.test(HypMut$tumor_var_freq,HypMut$hypoxia_score_buffa, method = 'spearman') # very positive
Agg = aggregate(list(HypMut$hypoxia_score_buffa,HypMut$tumor_var_freq), by = list(HypMut$sample,HypMut$Tier2), FUN = median)
names(Agg) = c('sample','Tier2','MedianHypoxiaScoreBuffa','MedianVaf')
cor.test(Agg$MedianHypoxiaScoreBuffa,Agg$MedianVaf, method = 'spearman') # still positive

##### 3B:  if there is correlation between VAF and hypoxia within numerous cancer types? (the strongest correlation (minimal p) is in the most numerous kidney)
names(HypMut)
Patients = HypMut[,c(1,3,10,11)]; Patients = unique(Patients); nrow(Patients) # 828
Tissues = data.frame(table(Patients$Tier2)) # 19 vs 21?
Tissues = Tissues[order(-Tissues$Freq),]
NumerousTissues = Tissues[Tissues$Freq >= 50,]$Var1; length(NumerousTissues) 
NumerousTissues # Breast   Kidney   Liver    Lung     Ovary    Pancreas

cor.test(Agg[Agg$Tier2 == NumerousTissues[1],]$MedianHypoxiaScoreBuffa,Agg[Agg$Tier2 == NumerousTissues[1],]$MedianVaf, method = 'spearman') # 
cor.test(Agg[Agg$Tier2 == NumerousTissues[2],]$MedianHypoxiaScoreBuffa,Agg[Agg$Tier2 == NumerousTissues[2],]$MedianVaf, method = 'spearman') #  !!!
cor.test(Agg[Agg$Tier2 == NumerousTissues[3],]$MedianHypoxiaScoreBuffa,Agg[Agg$Tier2 == NumerousTissues[3],]$MedianVaf, method = 'spearman') #  !!!
cor.test(Agg[Agg$Tier2 == NumerousTissues[4],]$MedianHypoxiaScoreBuffa,Agg[Agg$Tier2 == NumerousTissues[4],]$MedianVaf, method = 'spearman') #
cor.test(Agg[Agg$Tier2 == NumerousTissues[5],]$MedianHypoxiaScoreBuffa,Agg[Agg$Tier2 == NumerousTissues[5],]$MedianVaf, method = 'spearman') #
cor.test(Agg[Agg$Tier2 == NumerousTissues[6],]$MedianHypoxiaScoreBuffa,Agg[Agg$Tier2 == NumerousTissues[6],]$MedianVaf, method = 'spearman') # 

##### 4:  if there is correlation between A>G and hypoxia (Whole dataset and within numerous cancer types)?
##### 5:  if there is correlation between Absolute number of mtDNA mutationsand hypoxia (Whole dataset and within numerous cancer types)?

names(HypMut)
Patients = HypMut[,c(1,3,10,11)]; Patients = unique(Patients); nrow(Patients) # 828
Tissues = data.frame(table(Patients$Tier2)) # 19 vs 21?
NumerousTissues = Tissues[Tissues$Freq >= 50,]$Var1; length(NumerousTissues) 
NumerousTissues # Breast   Kidney   Liver    Lung     Ovary    Pancreas

i = 1
temp = HypMut[HypMut$Tier2 == NumerousTissues[i],]
summary(glm(temp$AhGhDummy ~ temp$hypoxia_score_buffa, family = binomial()))










#################################################
####### OLD CODE:
#################################################


## AhGhfr is expected to be higher among high VAF and lower among hypoxic

TvVec = c('A>T','A>C','C>A','C>G','T>A','T>G','G>C','G>T')

table(HypMut$Subst) # light chain
str(HypMut$tumor_var_freq)
HypMut$AhGhDummy = 0
for (i in 1:nrow(HypMut)) {if(HypMut$Subst[i] == 'T>C') {HypMut$AhGhDummy[i] = 1}}
table(HypMut$AhGhDummy)
summary(glm(HypMut$AhGhDummy ~ HypMut$hypoxia_score_ragnum + HypMut$tumor_var_freq, family = binomial()))
summary(glm(HypMut$AhGhDummy ~ HypMut$tumor_var_freq, family = binomial())) # a bit

### frequencies of all four transitions positively correlate with hypoxic score (the higher the score => the higher VAF => early origin and/or more relaxed mtDNA selection in hypoxic cancers)
summary(lm(HypMut$hypoxia_score_ragnum ~ HypMut$tumor_var_freq)) # very positive => the higher the hypoxia the higher VAF (the older all mutations => originated at healthy tissues)
summary(lm(HypMut[HypMut$Subst == 'T>C',]$hypoxia_score_ragnum ~ HypMut[HypMut$Subst == 'T>C',]$tumor_var_freq)) # positive
summary(lm(HypMut[HypMut$Subst == 'C>T',]$hypoxia_score_ragnum ~ HypMut[HypMut$Subst == 'C>T',]$tumor_var_freq)) # positive
summary(lm(HypMut[HypMut$Subst == 'G>A',]$hypoxia_score_ragnum ~ HypMut[HypMut$Subst == 'G>A',]$tumor_var_freq)) # positive
summary(lm(HypMut[HypMut$Subst == 'A>G',]$hypoxia_score_ragnum ~ HypMut[HypMut$Subst == 'A>G',]$tumor_var_freq)) # positive
summary(lm(HypMut[HypMut$Subst %in% TvVec,]$hypoxia_score_ragnum ~ HypMut[HypMut$Subst  %in% TvVec,]$tumor_var_freq)) # positive

# emerging PolG signature? check it!

## can we associate hypoxia with cell division rate of each tissue - YES (correlation is weak, but expected direction)
# from Cancer.DifferencesBetweenCancerTypes.R


# analyses


# if we repeat the same with rare substitutions only - effect is even better - why?
summary(HypMut$tumor_var_freq) # 5.3700
Agg = aggregate(HypMut[HypMut$tumor_var_freq <1.79,]$hypoxia_score_buffa, by = list(HypMut[HypMut$tumor_var_freq <1.79,]$Tier2,HypMut[HypMut$tumor_var_freq <1.79,]$TurnOverDays), FUN = median)
names(Agg) = c('CancerTissue','TurnOverDays','MedianHypoxiaScoreBuffa')
plot(Agg$TurnOverDays,Agg$MedianHypoxiaScoreBuffa)
cor.test(Agg$TurnOverDays,Agg$MedianHypoxiaScoreBuffa, method = 'spearman', alternative = 'less') # 0.02704


## can we rerun the same lm where instead of T>C there is a hypoxia
summary(lm(HypMut$hypoxia_score_buffa ~ HypMut$tumor_var_freq+HypMut$TurnOverDays)) # positive both coefficients!!!
summary(lm(HypMut$hypoxia_score_buffa ~ 0 +  HypMut$tumor_var_freq+HypMut$TurnOverDays)) # positive both coefficients!!!

Final$OtherMut = 1-Final$AhGhfr
colors = c("red","black")

FinalNew = data.frame(Final$AhGhfr, Final$OtherMut, rownames(FinalNew), Final$hypoxia_score_buffa)    
FinalHypo = FinalNew[FinalNew$Final.hypoxia_score_buffa>=32,]
FinalHypo$Final.hypoxia_score_buffa = NULL
FinalNorm = FinalNew[FinalNew$Final.hypoxia_score_buffa<32,]
FinalNorm$Final.hypoxia_score_buffa = NULL

a = FinalHypo %>% gather(subtype, freq,  Final.AhGhfr:Final.OtherMut)

b = FinalNorm %>% gather(subtype, freq,  Final.AhGhfr:Final.OtherMut)

pdf("../../Body/4Figures/Cancer.Hypoxia.pdf", width=7, height=3)
ggplot(a, aes(fill=subtype, y=freq, x=rownames.FinalNew.)) + 
  geom_bar(position="fill", stat="identity")+scale_fill_manual(values=c("red","black"))

ggplot(b, aes(fill=subtype, y=freq, x=rownames.FinalNew.)) + 
  geom_bar(position="fill", stat="identity")+scale_fill_manual(values=c("red","black"))
dev.off()

